# January 12, Learning R

```{r knitr learningR, include = FALSE}
knitr::opts_knit$set(message = FALSE, warning = FALSE)
```

```{r load packages, include=FALSE}
library(tidyverse)
library(socviz)
library(gapminder) # data on life expectancy, GDP per capita, pop 
library(here)
library(palmerpenguins)
```

### What R looks like

Code you can type and run

```{r}
# Inside code chunks, lines beginning with a # character are comments
# Coments are ignored by R

my_numbers <- c(1, 1, 2, 4, 1, 3, 1, 5)
```

By convention, code output in documents is prefixed by \#\#

```{r code, echo = FALSE}
my_numbers
```

### Everything in R has a name

```{r names}
my_numbers
letters
pi
```

**Forbidden** names

```{r forbbidden names, eval = FALSE}
TRUE
FALSE
Inf
NaN
NA

for
while
if
```

The `c()` function *combines* or *concatenates* things

```{r c function}
object <- c("b", "e", "n")
object
```

Use alt and - on Windows to use the assignment operator `<-` inside an code chunk. On the other hand, `=` has a different meaning when used in functions. You can use it for the `<-` assignment though. Just be consistent.

```{r object}
object = c("b", "e", "n")
object
```

### Do things to objects with a function

Functions take *inputs* to *arguments*. They perform *actions*. They produce, or return, *outputs*

```{r function}
mean(x = my_numbers)
```

In this case, the input is `my_numbers`, the argument is the vector `x`. The action is calculating the mean (e.g. the sum of the vector over its length). Finally, the function returns a vector of 1 length as an output, in other words, the mean of 2.25.

When you do not know what a function does, use `?` or `??` in the console followed by the function's name to read its documentation.

Functions can return things besides vectors.

```{r function outputs}
summary_numbers <- summary(my_numbers)
summary_numbers
class(summary_numbers)
```

The `rm` command can get rid of objects in the environment.

```{r rm}
rm(summary_numbers)
```

Functions can also be nested.

```{r nested functions}
c(1:20)

mean(c(1:20))

summary(mean(c(1:20)))

names(summary(mean(c(1:20))))

length(names(summary(c(1:20))))
```

Instead of nesting functions in parentheses, we can use the pipe operator `%>%`. This places the previous output into the first argument of the next. We can typically read the pipe as "and then."

```{r pipe}
c(1:20) %>% 
  mean() %>% 
  summary() %>%
  names() %>% 
  length()
```

The shortcut for the pipe operator is control + shift + m.

### Types and classes

The object inspector in RStudio is your friend.

You can ask an object what it is.

```{r}
class(letters)

typeof(letters)
```

It is important to know what class or type an object because functions may only work with certain types.

```{r}
sum(my_numbers)

# sum(letters) would not work
```

### First plots

```{r first plot}
p <- ggplot(data = gapminder,
            mapping = aes(x = log(gdpPercap),
                          y = lifeExp))
p + geom_point()

# the same plot could be created using geom smooth

p + geom_smooth()
```
