# Making Graphs

```{r knitr graphs, include = FALSE}
knitr::opts_knit$set(message = FALSE, warning = FALSE)
```

```{r load packages graphs, include=FALSE}
library(tidyverse)
library(socviz)
library(gapminder) # data on life expectancy, GDP per capita, pop 
library(here)
library(palmerpenguins)
```

> The peer review process... and the entire scientific enterprise is not designed to detect fraud.

It is hard to design proceedings to ensure that people are honest with their publications. We often find problems after the fact.

-   Ex: Honor Code

We write out instructions or recipes, rather than doing a series of point-and-click steps.

```{r create plot}
p <- ggplot(data = gapminder, 
            mapping = aes(x = gdpPercap,
                          y = lifeExp)) # links variables to graphical elements on the plot.
p + geom_point()

```

In R, we create **objects** by assigning something to a name. The **assignment operator** (\<-) attaches the output of the function `ggplot` to a the name `p`.

Other aesthetic mappings can include, e.g., `color`, `shape`, and `size`.

-   Mappings do not directly specify the particular, e.g., colors, shapes, and sizes that will appear on the plot. Instead they establish which variables will be represented by which visible elements.

### What exactly is a graph anyway?

We encode the data to create a "faithful representation" of the processes we see at play.

The way people interpret plots depends on how well the data visualization can track back to the data.

```{r second graph}
final_plot_EDIT1 <- p + geom_point() + 
  geom_smooth(method = 'lm') + # The geoms are drawn in the order that you provide them
  scale_x_log10(label = scales::dollar_format()) + # Transforms x scale to a log scale. Does not change the data.
  labs(x = 'GDP Per Capita',
       y = 'Life Expectancy in Years',
       title = 'Economic Growth and Life Expectancy',
       subtitle = 'Data points are country-years',
       caption = 'Data source: Gapminder')

final_plot_EDIT1
```

The process of data visualization is additive. We supply functions to change the graph.

-   `scales` package provides graphical scales to map data to aesthetics. Automatically determine breaks and labels.

### How ggplot works: The flow of action

1.  Tidy data

2.  Mapping

3.  Geom

4.  Coordinates, scales

    1.  usual `coord_cartesian()` but sometime another (e.g. polar coordinates in W.E.B. Du Bois plots from World Exposition)

5.  Labels and guides

    1.  `labs()` and `guides()`

6.  Themes

    1.  `theme_minimal()`

There are two kinds of elements:

1.  Visual elements that logically represent the plot (e.g. x and y axis representing quantities)
2.  Others that are not connected to the data set and strictly thematic. These are controlled by a separate set of functions.

## January 26, Making Graphs (Continuation)

### Mapping vs. setting aesthetics

-   Mapping allows the data to be represented. Setting provides a certain visual property that is not dependent on data.

-   Mapping occurs in `aes()` function. Setting occurs in geom functions.

```{r mapping vs setting}
# color is mapped incorrectly. The property of color is mapped to the character string 'purple,' not the color of purple.
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp,
                          color = 'purple')) 

# color is set correctly.
p + geom_point() +
  geom_smooth(method = 'loess') +
  scale_x_log10()

p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp))

p + geom_point(color = 'purple') +
  geom_smooth(method = 'loess') +
  scale_x_log10()
```

In data visualization, some elements will be mapped to **variables**, while others will be set to **values**.

The mapping is passed onto the succeeding geoms.

```{r mapping}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp,
                          color = continent,
                          fill = continent))

# Very "busy" graph.
p + geom_point() +
  geom_smooth(method = 'loess') +
  scale_x_log10()

```

Pay close attention to which scales and guides are drawn, and which aren't.

-   The guide shows the color of the point, line, and fill.

-   But we can change by moving the mapping to geom functions rather than `ggplot()`.

```{r altering mapping}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp))

# the geom_smooth function does not alter based on continent
p + geom_point(mapping = aes(color = continent)) +
  geom_smooth(method = 'loess') +
  scale_x_log10()
```

Every mapped variable has a scale. Scales are represented by guides by default, unless explicitly removed.

### Palmer Penguins and Simpson's Paradox

-   Changing the mapping of certain aesthetics among geoms can change the interpretation of the graph.

```{r simpsons paradox}
p <- ggplot(data = penguins,
            mapping = aes(x = bill_length_mm,
                          y = bill_depth_mm))

p + geom_point(mapping = aes(color = species)) +
  geom_smooth(mapping = aes(color = species, fill = species), method = 'lm', se = FALSE) +
  geom_smooth(method = 'lm', se = FALSE) +
  labs(x = "Bill length (mm)",
       y = "Bill depth (mm)",
       title = "Simpson's Paradox among Penguin Dimensions",
       subtitle = "After Separating by Species",
       caption = "Source: Palmer Station Long Term Ecological Research") 
  theme_bw()
```

> Like other rules of syntax, the grammar limits what you can validly say, but it doesn't make what you say intelligible or meaningful.

### Grouped data and group aesthetic

```{r grouped data}
p <- ggplot(data = gapminder,
            mapping = aes(x = year,
                          y = gdpPercap))

# the graph looks bad because the data has more structure than the graph knows. The unit is country-year, not year.
p + geom_line()

# now adding a grouped aesthetic for country.
p <- p + geom_line(mapping = aes(group = country))

# facet wrapping is not a geom. It is a way of arranging geoms.
p + facet_wrap(~ continent)

p + geom_line(color = "gray70",
              mapping = aes(group = country)) +
  geom_smooth(size = 1.1,
              method = 'loess',
              se = FALSE) +
  scale_y_log10(labels = scales::dollar) +
  facet_wrap(~ continent, ncol = 5) +
  labs(x = "Year",
       y = "GDP per Capita",
       title = "GDP per capita on Five Continents")

```
