# Textual Analysis

## Gutenberg data

```{r knitr textual analysis, include = FALSE}
knitr::opts_chunk$set(message = FALSE)
```

```{r libraries textual analysis, include = FALSE}
library(gutenbergr)
library(tidytext)
library(socviz)
library(tidyverse)
library(gapminder)
```

Gutenberg package helps us download plain text files of literature included in the repository.

```{r}
# Finding author
gutenberg_works() %>%
  filter(gutenberg_author_id == 1039)

# Downloading Ulysses
raw_text <- gutenberg_download(gutenberg_id = 4300)

# Tidying data, tokenization process removes possessives, punctuation.
full_text <- raw_text %>% 
  mutate(line = row_number()) %>% 
  unnest_tokens(output = word, input = text, token = 'words')
```

```{r}
full_text %>% 
  count(word, sort = TRUE)

head(stop_words)
```

Stop words include the, and, of, and other nuts and bolts of grammar.

We will anti-join by taking out any words in `full_text` that are included in `stop_words`.

```{r}
full_text %>%
  anti_join(stop_words) %>%
  filter(! str_detect(word, "'")) %>%
  filter(! str_detect(word, "’")) %>%
  count(word, sort = TRUE) %>%
  top_n(20) %>%
  mutate(word=reorder(word, n))
```

What sort of questions can we ask about this data? One option is sentiment analysis, a means of determining whether the text is more positive or negative.

```{r}
full_text %>% 
  inner_join(get_sentiments('bing')) %>% 
  count(sentiment, word, sort = TRUE) %>% 
  group_by(sentiment) %>% 
  arrange(desc(n)) %>% 
  slice(1:20) %>% 
  ungroup() %>% 
  mutate(word = reorder(word, n)) %>% 
  ggplot(mapping = aes(x = n, y = word, fill = sentiment)) +
  geom_col() +
  facet_wrap(~ sentiment, scales = "free") +
  guides(fill = 'none')
```


```{r}
full_text %>% 
  inner_join(get_sentiments('nrc')) %>% 
  count(sentiment, word, sort = TRUE) %>% 
  group_by(sentiment) %>% 
  slice(1:10) %>% 
  ungroup() %>% 
  mutate(word = reorder(word, n)) %>% 
  ggplot(mapping = aes(x = n,
                       y = word,
                       fill = sentiment)) +
  geom_col() +
  guides(fill = 'none') +
  facet_wrap(~ sentiment,
             scales = 'free',
             ncol = 3)
```

Now with Bigrams

```{r}
tidy_ngram <- raw_text %>% 
  unnest_tokens(output = bigram, input = text, token = 'ngrams', n = 2) %>% 
  separate(bigram, c('word1', 'word2'), sep = ' ') %>% 
  filter(!word1 %in% stop_words$word, !word2 %in% stop_words$word) %>% 
  count(word1, word2)
```

```{r}
he_she <- raw_text %>% 
  unnest_tokens(bigram, text, token = 'ngrams', n = 2) %>% 
  separate(bigram, c('word1', 'word2'), sep = ' ') %>% 
  filter(word1 %in% c('he', 'she', 'they')) %>% 
  filter(!word2 %in% stop_words$word, !str_detect('word2', "'")) %>% 
  count(word1, word2)

he_she %>% 
  group_by(word1) %>% 
  slice(1:20) %>% 
  ggplot(mapping = aes(x = n, y = word2, fill = word1)) +
  geom_col() +
  guides(fill = 'none') +
  facet_wrap(~ word1, scales = 'free')
```


## Tidydata

Underpinning any textual analysis is *tidy* data (tibble).

- Every variable gets its own column
- Every row is an observation
- Every cell is a value.

`ggplot` accepts tidy data. Some data frames are not tidy.

- Example: Census data can organize education by several variables, such as elementary education being split into 0-4 years and 5-8 years. Spreading a variable across several columns means that it isn't tidy.

- Excel sheets with a title row going across the data. Bolded or shaded information in a column that denotes a certain variable.

Pivot commands can transform the shape of the data. Usually we need to move to longer format.

```{r}
head(edu, 5) # Needs to be transformed into long-form

# Pick the columns that need to be put into a single column in cols argument. Then name the column. Finally, assign the values for the four colums to value_to argument.

edu %>% 
  pivot_longer(cols = elem4:coll4, names_to = "educ", values_to = "n") %>% 
  head(5)
```

Pivoting wider

```{r}
# Take every value from year into its own column

gapminder %>%  
  pivot_wider(names_from = year, values_from = lifeExp:gdpPercap) %>% 
  head(5)
```

## Saving your Work

`ggsave` helps save individual plots. By default, it saves the most recent plot that was made. More easily, `ggsave` can save objects in the environment.

