# Tidy Data

```{r knitr tidy, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

```{r setup tidy, include=FALSE}
library(tidyverse)
library(socviz)

theme_set(theme_minimal())
```

> Problem set 2 will substitute the midterm project.

We want to transform and summarize first before sending clean tables to `ggplot()`.

## Counts are easy

Use `geom_bar` to summarize counts. Stacked bar charts are not easy to interpret, so we can change `position = 'dodge'`, we can see the proportions of counts using the `mapping = fill`.

## Transform data first if dealing with proportions or other summary statistics

Row percents/row marginals and column percents/columns marginals (do rows sum to 100% or do columns?). We can also calculate total percents.

`dplyr` lets you manipulate tables in a series of steps or a **pipeline**.

### `group_by()`

Groups the data at a lavel we want, such as "religion by Region" or "Authors by Publications by Year".

### `filter()` and `select()`

Subsets table based on rows and columns, respectively

### `Mutate()`

changes the data by creating new cvariables at the current level of grouping.

### `Summarize()`

Creates summary statistics for grouped data.

## GSS Data example

```{r gss longhand}
rel_by_region <- gss_sm %>% 
  group_by(bigregion, religion) %>% # group by bigregion and religion. Read right to left using 'within' or left to right using 'by'
  summarize(n = n()) %>% 
  mutate(freq = n / sum(n),
         pct = round((freq * 100), 1))
```

### Objects in pipeline carry forward assumptions

For example, the `group_by()` groups carries over to the `summarize()` and `mutate()` functions.

The variable `n` created by `summarize()` can be used in the `mutate()` line. We can also use `freq` immediately after creating it.

### Shorthand version

```{r gss shorthand, echo=FALSE}
gss_sm %>% 
  group_by(bigregion, religion) %>% 
  tally() # Produces the same output as summarize without as much code.

gss_sm %>% 
  count(bigregion, religion) # No group_by necessary, but this means no groups are created.
```

### Pipeline tables are easier to check for errors

```{r pipeline easy, echo=FALSE}
rel_by_region %>% 
  group_by(bigregion) %>% 
  summarize(total = sum(pct))
```

### GSS Graph

```{r gss plot}
p <- ggplot(data = rel_by_region,
            mapping = aes(x = bigregion,
                          y = pct,
                          fill = religion))

p + geom_col(position = 'dodge')
```

But this is not an effective graph!

```{r}
p <- ggplot(data = rel_by_region,
            mapping = aes(x = pct,
                          y = religion,
                          fill = religion))

p + geom_col(position = 'dodge') +
  labs(x = 'Percent', y = NULL) +
  guides(fill = 'none') +
  facet_wrap(~ bigregion, nrow = 1)
```

- Try putting categories on the y-axis. 

- Try faceting variables instead of mapping them.

- Try to minimize the need for guides and legends.

> Note that these rules are not absolute.

## Other forms of facetting

### Facetting multiple variables

```{r facet wrap multiple}
p <- ggplot(data = gss_sm, mapping = aes(x = age, y = childs))

p + geom_point(alpha = 0.2) +
  geom_smooth() + facet_wrap(~ sex + race, nrow = 1) + # The formula syntax isn't limited to one variable
  theme_minimal()
```

### Facet grid

```{r facet grid}
p + geom_point(alpha = 0.2) +
  geom_smooth() + facet_grid(race ~ sex) # The dimensions of the table are defined. Race by sex.
```

Be careful of breaking the data into too many categories; data may not have enough observations of each category.

## The Process

- Full data and then

- Subset data and then

- Grouped calculation and then

- New columns or further summary and then

- Plot and polish.

## Organ Data

```{r organ plot 1}
organdata %>% 
  ggplot(mapping = aes(x = year, y = donors, group = country)) +
  geom_line() + facet_wrap(~ reorder(country, -donors, FUN = mean, na.rm = TRUE))
```

```{r organ plot 2}
organdata %>% 
  ggplot(mapping = aes(x = donors, y = reorder(country, donors, fill = world, na.rm = TRUE))) +
  geom_boxplot() + 
  labs(x = "Donors", y = NULL, fill = "Welfare State")
```

