# Pandemic Data

```{r setup pandemic, include = FALSE, messag=FALSE}
library(tidyverse)
library(socviz)
library(here)
library(covdata)
library(slider)
library(patchwork)
```

```{r knitr pandemic, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, collapse=TRUE)
```

## Vegetable Fat and Animal Fat Plot

The amount of vegetable fat is increasing but animal fat mostly steady. The conclusion is that "added fats and oils... increased by almost 30 pounds between 1970 and 2010." 

- The number of firms reporting vegetable oil production increased.

- **After a sudden change in a variable over time, the question should be "how has the collection of this data changed?"**

## COVID-19

Similarly, data collection about COVID-19 has changed over time, distorting the statistics that we might use.

Reporting delays alter interpretations. 

- So date of report vs. date of specimen collection are altered by weekly trends (more tests on Monday). Inferences can lead analysts to conclude the cases/tests are falling when they aren't.

```{r pandemic data}
country_rank <- stmf %>% 
  filter(sex == 'b', year > 2014 & year < 2020) %>% 
  select(country_code, cname, iso3, year:sex, age_group, death_rate, rate_total) %>% 
  group_by(country_code) %>% 
  summarize(mean_rate = mean(rate_total, na.rm = TRUE)) 
```

### `dplyr`'s window functions

```{r cumsum}
covnat_weekly %>% 
  filter(iso3 == 'USA') %>% 
  select(date, cname, iso3, cases) %>% 
  mutate(cumulative = cumsum(cases))
```

`cume_dist()` gives the proportion of values less than or equal to the current

```{r cume_dist}
covnat_weekly %>% 
  select(date, cname, iso3, deaths) %>% 
  filter(iso3 == 'USA') %>% 
  mutate(cume_dist(desc(deaths) < 0.1)) # i.e. top ten percent
```

`lead()` and `lag()` functions allow you to access the previous and next values in a vector

```{r lag}
my_vec <- c(1:20) 
my_vec

lag(my_vec)

my_vec - lag(my_vec)
```

So this allows us to calculate off-sets.

```{r covus}
# covus %>% 
#   select(-date_quality_grade) %>% 
#   filter(measure == 'death') %>% 
#   group_by(state) %>% 
#   arrange(date) %>% 
#   mutate(deaths_daily = count - lag(count, order_by = date))
```

We can always write functions using `function()`.

```{r get_daily_count}
get_daily_count <- function(count, date) {
  count = lag(count, order_by = date)
}
```

It is more useful to use functions when we want to perform several steps on the data.

`dplyr`'s window functions don't include moving averages, so we can use the `slider` package.

```{r slider}
covus %>%
  select(-data_quality_grade) %>%
  filter(measure == 'death') %>%
  group_by(state) %>%
  arrange(date) %>%
  mutate(deaths_daily = get_daily_count(count, date),
         deaths7 = slide_mean(deaths_daily, before = 7, na_rm = TRUE)) %>%
  arrange(state, desc(date)) %>%
  filter(state %in% 'NY')
```
There are other `slider` functions like `slide_max()` and `slide_min()`.

### Patchwork

`patchwork` allows us to arrange several plots together.

```{r pathwork}
# p1 + p2 + p3 + guide_area() +
# plot_layout(ncol = 2, guides = 'collect')
```

`guides = 'collect'` combines legends together so that it doesn't appear multiple times.

## Organ Data

```{r geom jitter}
p <- ggplot(data = organdata,
            mapping = aes(x = donors,
                          y = reorder(country, donors, na.rm = TRUE),
                          color = world)) +
  geom_jitter() + labs(x = NULL) +
  theme(legend.position = "top")

p1 <- p + geom_jitter(position = position_jitter(width = 0.15)) + 
  labs(x = NULL) + theme(legend.position = "top")

p + p1
```

```{r by country}
by_country <- organdata %>% group_by(consent_law, country) %>%
    summarize(donors_mean= mean(donors, na.rm = TRUE),
              donors_sd = sd(donors, na.rm = TRUE),
              gdp_mean = mean(gdp, na.rm = TRUE),
              health_mean = mean(health, na.rm = TRUE),
              roads_mean = mean(roads, na.rm = TRUE),
              cerebvas_mean = mean(cerebvas, na.rm = TRUE))


# summarize across several variables

by_country <- organdata %>% 
  group_by(consent_law, country) %>%
    summarize(across(where(is.numeric),
                     list(mean = mean, 
                          sd = sd),
                      na.rm = TRUE,
                      .names = "{col}_{fn}"),
              .groups = "drop")
```

```{r country means}
p <- ggplot(data = by_country,
            mapping = aes(x = donors_mean,
                          y = reorder(country, donors_mean),
                          color = consent_law))

p + geom_point(size = 3) +
  labs(x = 'Donor Procurement Rate',
       y = NULL, color = 'Consent Law') +
  theme(legend.position = 'top')
```

```{r facet wrap organdata}
p <- ggplot(data = by_country,
            mapping = aes(x = donors_mean, y = reorder(country, donors_mean, na.rm = TRUE)))

p + geom_point(size = 3) +
  facet_wrap(~ consent_law, scales = 'free_y', ncol = 1) +
  labs(x = 'Donor Procurement Rate',
       y = '')
```
Add ranges.

```{r geom_pointrange}
p <- ggplot(data = by_country, 
            mapping = aes(x = reorder(country, donors_mean, na.rm = TRUE),
                          y = donors_mean))

p + geom_pointrange(mapping = aes(ymin = donors_mean - donors_sd, 
                                  ymax = donors_mean + donors_sd)) +
  labs(y = 'Donor Procurement Rate',
       x = '') +
  coord_flip()
```

```{r elections_historic}
library(ggrepel)

elections_historic %>% select(2:7)
```

```{r president plot}
p_title <- 'Presidential Elections: Popular & Electoral College Margins'
p_subtitle <- '1824-2016'
p_caption <- 'Data for 2016 are provisional'
x_label <- "Winner's share of Popular Vote"
y_label <- "Winner's share of Electoral College Votes"

p <- ggplot(data = elections_historic, aes(x = popular_pct, y = ec_pct,
                                           label = winner_label))

p + geom_hline(yintercept = 0.5, size = 1.4, color = "gray80") +
  geom_vline(xintercept = 0.5, size = 1.4, color = "gray80") +
  geom_point() +
  geom_text_repel() +
  scale_x_continuous(labels = scales::percent) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = x_label, y = y_label, title = p_title, subtitle = p_subtitle,
       caption = p_caption)
```

```{r selective labels}
p <- ggplot(data = by_country,
            mapping = aes(x = gdp_mean, y = health_mean))

p + geom_point() +
  geom_text_repel(data = subset(by_country, gdp_mean > 25000),
                  mapping = aes(label = country))

p <- ggplot(data = organdata,
            mapping = aes(x = roads, y = donors))

p + geom_point() +
  annotate(geom = "rect", xmin = 125, xmax = 155,
           ymin = 30, ymax = 35, fill = "red", alpha = 0.2) +
  annotate(geom = "text", x = 157, y = 33,
           label = "A surprisingly high \n recovery rate", hjust = 0)
```

