# Missing Data

```{r knir missingness, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

There are two problems of missing data in data-- selection bias and missingness.

## Selection Bias

Survivorship bias: World War II planes coming back with bullet holes. The places that there weren't holes possibly led to more downed planes.

- The missingness of some data is salient (i.e. representative of bias, data collection is systematically biased).

- Development of statistics is related to development of state (war, market regulation, legal system).

  - Reintegration into civilian life.
  
Other forms of selection bias:

- Answering a poll emailed by a Republican-affiliated group/candidate, voting on a website.

- Self-help books and looking only at "successful" or positive outcomes. Those that experience negative outcomes may have been doing the same thing.

### Simpson's Paradox

A variety of selection bias that is produced from omitting a key variable.

- Ex: Palmer's penguins and omission of sex or species.

- Berkley's admission case: Women weren't necessarily discriminated against, they applied to more competitive programs.

> How to produce plots showed in slides?

### Collider Bias

When we control for an additional variable that is actually important (confounding).

- Ex: Talent vs. looks in Hollywood vs. the general population.

```{r library missingness, include = FALSE}
library(tidyverse)
library(socviz)
library(naniar)
library(visdat)

theme_set(cowplot::theme_cowplot())
```

### Hollywood

```{r hollywood}

df <- tibble(looks = rnorm(1000),
             talent = rnorm(1000),
             total = looks + talent,
             hollywood = total > quantile(total, 0.99)) # hollywood is top 1%

df %>% 
  ggplot(mapping = aes(x = talent,
                       y = looks,
                       color = hollywood)) +
  geom_point(alpha = 0.4) + 
  geom_smooth(method = "lm", se = F) +
  scale_color_manual(values = c("navy", "yellow2")) +
  labs(x = "Talent", y = "Looks", color = "Hollywood") +
  theme(legend.position = "bottom")
```

## Missing Observations

### Organ data and `visdat`

```{r organdata missing}
vis_dat(organdata)
miss_var_summary(organdata) # Which variables have more missing data
miss_case_summary(organdata) # Which cases have the most NAs for variables
vis_miss(organdata, cluster = TRUE)
gg_miss_upset(organdata) # Relationships in missingness between several variables

ggplot(organdata,
       aes(x = donors,
           y = pubhealth)) +
  geom_miss_point()
```

It's important to look at the missingness of data before fitting a model. 

### Congressional Data

#### Upset Plots

