# Making Maps

Geography was deinstitutionalized as a social science in the United States around the 1920's.

- Connection to social geography, anthropology, and implementation in colonial projects.

- So geography acted as a glue for emerging social sciences even after its decline.

- 19th century theory connected land to culture (often to racialize culture).

## Choropleths

Show regions that are shaded or colored, cross-hashed by a quantity of interest.

- Ex: electoral maps

Fundamental problem is that the space/size can be confused as a significant characteristic of the data when it really isn't (e.g. the electoral victor of the most area may not have the same proportion of votes).

- Correct by manipulating the physical geography by the population or some other variable.

A cartogram turns geography into blocks or a series of polygons (e.g. electoral maps that have boxes or hexagons).

- Looks less warped but still kind of weird.

## State-level Election Data

```{r maps load libraries, include = FALSE, message = FALSE, warning=FALSE}
library(tidyverse)
library(socviz)
library(colorspace)
library(sf)
library(nycdogs)
library(tigris)
library(tidycensus)
# library(cartogram)
```

```{r maps knitr, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r theme negate, include = FALSE}
`%nin%` = Negate(`%in%`)

theme_map <- function(base_size = 9, base_family = "") {
  require(grid)
  theme_bw(base_size=base_size, base_family=base_family) %+replace%
    theme(axis.line=element_blank(),
              axis.text=element_blank(),
              axis.ticks=element_blank(),
              axis.title=element_blank(),
              panel.background=element_blank(),
              panel.border=element_blank(),
              panel.grid=element_blank(),
              panel.spacing=unit(0, "lines"),
              plot.background=element_blank(),
              legend.justification = c(0,0),
              legend.position = c(0,0))
}

theme_nymap <- function(base_size=9, base_family="") {
    require(grid)
    theme_bw(base_size=base_size, base_family=base_family) %+replace%
        theme(axis.line=element_blank(),
              axis.text=element_blank(),
              axis.ticks=element_blank(),
              axis.title=element_blank(),
              panel.background=element_blank(),
              panel.border=element_blank(),
              panel.grid=element_blank(),
              panel.spacing=unit(0, "lines"),
              plot.background=element_blank(),
              legend.justification = c(0,0),
              legend.position = c(0.1, 0.6), 
              legend.direction = "horizontal"
        )
}
```

```{r county pop map}
county_full <- left_join(county_map, county_data, by = 'id')

p <- ggplot(data = county_full,
            mapping = aes(x = long, y = lat,
                          fill = pop_dens,
                          group = group))

p1 <- p + geom_polygon(color = 'gray70', size = 0.1) + coord_equal()

p2 <- p1 + scale_fill_brewer(palette = "Blues",
                             labels = c("0-10", "10-50", "50-100", "100-500", "500-1,000", "1,000-5,000", ">5,000"))

p2 + labs(fill = "Population per\nsquare mile") + theme_map() +
  guides(fill = guide_legend(nrow = 1)) +
  theme(legend.position = "bottom")
```

### Using simple features and `geom_sf()`

```{r nycdogs data}
nyc_fb <- nyc_license %>% 
  group_by(zip_code, breed_rc) %>% 
  tally() %>% 
  mutate(freq = n / sum(n),
         pct = round(freq * 100), 2) %>%
  filter(breed_rc == "French Bulldog")

fb_map <- left_join(nyc_zips, nyc_fb)
```

```{r dog map}
fb_map %>% 
  filter(n > 1) %>% 
  ggplot(mapping = aes(fill = pct)) +
  geom_sf(color = "gray80", size = 0.1) +
  scale_fill_viridis_c(option = "A") +
  labs(fill = "Percent of All Licensed Dogs") +
  annotate(geom = "text", x = -74.145 + 0.029, y = 40.8075 - 0.012,
           label = "New York City's French Bulldogs", size = 5) +
  theme_nymap() +
  guides(fill = guide_legend(title.position = "top",
                             label.position = "bottom",
                             keywidth = 1, nrow = 1))
```

### Keeping Zero Count Rows

```{r license colors}
nyc_license %>% 
  filter(extract_year == 2018) %>% 
  group_by(zip_code, breed_rc) %>% 
  tally() %>% 
  mutate(freq = n / sum(n),
         pct = round(freq*100, 2)) %>% 
  filter(breed_rc == "French Bulldog")

nyc_fb <- nyc_license %>% 
  group_by(zip_code, breed_rc) %>% 
  tally() %>% 
  ungroup() %>% 
  complete(zip_code, breed_rc,
           fill = list(n = 0)) %>% 
  mutate(freq = n/sum(n),
         pct = round(freq*100, 2)) %>% 
  filter(breed_rc == "French Bulldog")

fb_map <- left_join(nyc_zips, nyc_fb)
```

```{r dog map colors}
fb_map %>% ggplot(mapping = aes(fill = pct)) +
    geom_sf(color = "gray80", size = 0.1) +
    scale_fill_continuous_sequential(palette = "Oranges") +
   labs(fill = "Percent of All Licensed Dogs in the City") +
  annotate(geom = "text", x = -74.145 + 0.029, y = 40.82-0.012, 
           label = "New York City's French Bulldogs", size = 6) + 
  annotate(geom = "text", x = -74.1468 + 0.029, y = 40.8075-0.012, 
           label = "By Zip Code. Based on Licensing Data", size = 5) + 
    theme_nymap() + 
    guides(fill = guide_legend(title.position = "top", 
                               label.position = "bottom",
                             keywidth = 1, nrow = 1))
```

## Population Components Example

```{r us_components}
not_lower48 <- c("Alaska", "Hawaii", "Puerto Rico")
  
us_components <- get_estimates(geography = "state", 
                               product = "components") %>% 
  filter(NAME %nin% not_lower48)

net_migration <- get_estimates(geography = "county",
                               variables = "RNETMIG",
                               year = 2019,
                               geometry = TRUE,
                               resolution = "20m") %>% 
  filter(str_detect(NAME, "Alaska") == F | !str_detect(NAME, "Hawaii") == F | !str_detect(NAME, "Puerto Rico") == F)
```

```{r migration wrangling}
order <- c("-15 and below", "-15 to -5", "-5 to +5", "+5 to +15", "+15 and up")

net_migration <- net_migration %>%
    mutate(groups = case_when(
      value > 15 ~ "+15 and up",
      value > 5 ~ "+5 to +15",
      value > -5 ~ "-5 to +5",
      value > -15 ~ "-15 to -5",
      TRUE ~ "-15 and below"
    )) %>%
    mutate(groups = factor(groups, levels = order)) 

 state_overlay <- states(
    cb = TRUE,
    resolution = "20m") %>%
    filter(GEOID != "72",
           NAME %nin% not_lower48)
```

```{r migration map}
ggplot() +
    geom_sf(data = net_migration, 
            mapping = aes(fill = groups, color = groups), 
            size = 0.1) +
    geom_sf(data = state_overlay, 
            fill = NA, color = "black", size = 0.1) +
    scale_fill_brewer(palette = "PuOr", direction = -1) +
    scale_color_brewer(palette = "PuOr", direction = -1, guide = "none") +
    coord_sf(datum = NA) +
    theme_minimal() +
    labs(title = "Net migration per 1000 residents by county",
         subtitle = "US Census Bureau 2019 Population Estimates",
         fill = "Rate",
         caption = "Data acquired with the R tidycensus package")
```
