# Maps and the Census

## Pitfalls with Spatial Data

Be careful of using zip codes, geographic entities where the values of focal variables simply restate population size, etc.

Feature of randomness is that it is clumpier than we expect.

- Ex: Plotting random points on a map.

- We should not assume that the clumpiness = structured process when it can also happen from random processes.

## NC Data

```{r maps census libraries, message= F, warning = F}
library(tidyverse)
library(tidycensus)
library(socviz)
library(sf)
library(babynames)
library(tigris)
```

```{r maps census knitr}
knitr::opts_chunk$set(message = F, warning = F)
```

```{r nc map}
# nc <- st_read(system.file("shape/nc.shp"), package = "sf")
```

```{r}
counties <- c("Wake County", "San Francisco County", "Durham County")

str_detect("[e]", counties)
```

## Baby Names

```{r}
popular_names <- babynames %>% 
  group_by(year, sex) %>% 
  slice_max(n = 3, order_by = prop)

babynames %>% 
  filter(sex == "F") %>% 
  group_by(year) %>% 
  mutate(popularity = -rank(prop)) %>% 
  filter(name == "Mary") %>% 
  ggplot(mapping = aes(x = year, y = popularity)) +
  geom_point()
```

```{r}
babynames %>% 
  filter(sex == "M" & name == "Oliver") %>% 
  group_by(year) %>% 
  ggplot(mapping = aes(x = year, y = prop)) +
  geom_line(size = 1.1) +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(breaks = seq(1880, 2015, by = 10)) +
  labs(y = "Percent of all Names", x = "Year", title = "Popularity of Oliver") +
  theme_light()
```

```{r most popular}
babynames %>% 
  group_by(year, sex) %>% 
  slice_max(n = 1, order_by = prop) %>% 
  ggplot(mapping = aes(x = year, y = prop, color = sex)) +
  geom_line(size = 1) + scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(breaks = seq(1880, 2015, by = 10)) +
  labs(y = "Percent", x = "Year",
       title = "Most Popular Name as a Percent of All Names",
       color = "Sex") +
  theme(legend.position = "top")
```

```{r unpopular names}
babynames %>% 
  group_by(year, sex) %>%
  filter(prop <= 0.001) %>%
  group_by(year, sex) %>% 
  tally() %>%   
  ggplot(mapping = aes(x = year, y = n, color = sex)) + 
  geom_line(size = 1) + scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = seq(1880, 2015, by = 10)) +
  labs(y = "Count", x = "Year", 
       title = "Names with a frequency of 1 in 1,000 or less",
       color = "Sex") + 
  theme(legend.position = "top")
```
