# Network Data

```{r network libraries, message=F, warning=F, include=F}
library(tidyverse)
library(socviz)
library(ggraph)
library(tidygraph)
library(kjhnet)
```

```{r network knitr, include=F}
knitr::opts_chunk$set(warning = F, message = F)
```

Networks are a kind of social structure. 

- **Nodes**: The individuals in a network connected by **edges**. 

- **Direct** ties consider whether relationships exist. **Indirect** considers the type of relationship between two nodes.

Simplest network is a diad (2 individuals with one link). 

Questions: How do we evaluate an edge/connection? 

- With people: do they have to be present, do they have to know each other at a certain level, etc.?

## Groups and Categories

A group is created relationaly (e.g. sorority member, friendship) whereas a category is determined based on non-relational criteria (e.g. being a woman, Christian).

### Group and ties

There is a duality between individuals who interact with others and the groups who interact with other groups.

Multiplying a matrix times its transposed matrix can create an estimation of the strength of an individual in a social network (i.e. how my organizations they are in as well as how many people).

Betweenness centrality vs. number of groups someone belongs. 

- Being regarded by others who are also well regarded.

## Ilyad

Direction of arrow: from victor to loser (direction of spear) or from loser to victor (direction of glory).

Degree centrality: The number of connections one has. In this context, the number of people one kills.

```{r il_tidy}
il_tidy

il_tidy %>% 
  activate(nodes) %>% 
  mutate(centrality = centrality_degree()) %>% 
  as_tibble() %>% 
  arrange(desc(centrality)) %>% 
  top_n(10, wt = centrality) %>% 
  ggplot(mapping = aes(x = centrality,
                       y = reorder(name, centrality),
                       color = affil)) +
  geom_point(size = 3) +
  labs(x = "Centrality", y = NULL, color = "Side") +
  theme_minimal() +
  theme(legend.position = "top")
```

```{r iliad graph}
label_colors <- c(prismatic::clr_lighten("blue", 0.7),
                  prismatic::clr_lighten("red", 0.7))
il_tidy %>%
  activate(nodes) %>%
  mutate(centrality = centrality_degree(mode = "out")) %>%
  ggraph(layout = "graphopt") +
  geom_edge_link(aes(start_cap = label_rect(node1.name),
                     end_cap = label_rect(node2.name)),
                 arrow = arrow(length = unit(1.5, 'mm'))) +
  geom_node_point(aes(color = affil)) +
  scale_color_manual(values = c("blue", "red"),
                     labels = c("Athenian", "Trojan")) +
  guides(color = "none", fill = "none") +
  geom_node_label(aes(filter = centrality > 0,
                      label = name, fill = affil),
                      size = rel(2.5)) +
  scale_fill_manual(values = label_colors) +
  labs(title = "Violence in The Iliad") +
  theme_graph() +
  theme(plot.title = element_text(size = rel(3)))
```

But the centrality of characters in the Ilyad not really determined by number of people they killed, but the importance of those that they did kill.

Betweenness centrality: The number of paths in a group that can go through a certain node.

- The latter emphasizes the relational structure of the Ilyad. 

```{r betweenness centrality}
il_tidy %>%
  activate(nodes) %>%
  mutate(centrality = centrality_degree(),
         betweenness = centrality_betweenness()) %>%
  as_tibble() %>%
  arrange(desc(betweenness)) %>%
  top_n(10, wt = betweenness) %>%
  ggplot(mapping = aes(x = betweenness,
                       y = reorder(name,
                       betweenness), color = affil)) +
  geom_point(size = 3) +
  labs(x = "Betweenness", y = NULL, color = "Side") +
  theme_minimal() +
  theme(legend.position = "top")
```

Alpha centrality: Evaluates centrality by the status captured by others, highly regarded from highly regarded (e.g. pecking order).

```{r alpha centrality}
il_tidy %>%
  activate(edges) %>%
  filter(act == "kills") %>%
  reroute(from = to, to = from) %>%
  activate(nodes) %>%
  mutate(alpha = centrality_alpha()) %>%
  as_tibble() %>%
arrange(desc(alpha)) %>%
    top_n(10, wt = alpha) %>%
  ggplot(mapping = aes(x = alpha,
                       y = reorder(name, alpha),
                       color = affil)) +
  geom_point(size = 3) +
  labs(x = "Alpha Centrality", y = NULL, color = "Side") +
  theme_minimal() +
  theme(legend.position = "top")
```


## Network relations in unexpected places

New professor exchanges by Ph.D. hires.

- It is helpful to transform individual data to relational to investigate structure even if it is not he final product.

## Paul Revere

### Matrix version

Matrix version of data

```{r head paul revere}
head(revere)
```

Where 0 indicates no membership and 1 represents membership

```{r revere matrix}
r_p <- as.matrix(revere[,-1]) %*% t(as.matrix(revere[,-1]))

# revere[,-1] represents all data except the first column.
# t() transposes, or flips the dimensions of a matrix (i.e. 2x3 becomes 3x2)

r_p[1:6, 1:6] # 

r_g <- t(as.matrix(revere[,-1])) %*% as.matrix(revere[,-1])

head(r_g)
```

### Tidy version

```{r revere_groups}
revere_groups

revere_groups %>% 
  ggraph(layout = "kk") +  # layout uses an algorithim to fill 2D space
  geom_edge_link(aes(width = value), color = "gray80") +
  geom_node_label(aes(label = name)) + theme_graph()
```

Tidy network data has two sets of data: edges and nodes.

- Edges in this lists the shared membership between any two groups using `value` column.

```{r revere centrality}
revere_persons 

revere_persons %>% 
  mutate(centrality = centrality_eigen()) %>% 
  ggraph(layout = "stress") +
  geom_edge_link0(aes(edge_width = value), color = "gray60") +
  scale_edge_width(range = c(0.02, 2)) +
  geom_node_point() +
  geom_node_label(aes(filter = centrality > 0.9, label = name), size = rel(2.5)) +
  theme_graph()
```

# The Iliad

