# Polishing your Plots

## Building a Plot a Piece at a Time

```{r polish libraries, message=F, warning=F}
library(socviz)
library(tidyverse)
library(ggrepel)
library(colorspace)
```

```{r}
head(asasec)

p <- ggplot(subset(asasec, Year == 2014),
            mapping = aes(x = Members, y = Revenues, label = Sname))

p + geom_point() + geom_smooth()
```

How to improve plot?

1. Improve smoother

2. Remove needless information like `se = T`

3. Add text 

4. Fix scales

5. Theme

```{r}
p + geom_smooth(method = 'lm', se = FALSE, color = 'gray80') +
  geom_point(mapping = aes(color = Journal)) +
  geom_text_repel(data = subset(asasec,
                                Year == 2014 & Revenues > 10000)) +
  scale_y_continuous(labels = scales::dollar_format())
```

In presentations, can use the layers to effectively communicate a story about the data.

- Communicated differently than when producing data visualizations for print.

## Theme functions

Built-in themes can be added per plot or globally set.

```{r include=FALSE}
theme_set(theme_bw()) # globally

p + theme_bw() # per plot
```

`theme()` function customizes individual elements in visualization.

- Should be used intentionally.

## Case studies

1. **Don't** use two y-axes with different scales.

  - Often used with time series. The use of two scales changes the relative shape of the distribution, can lead to misinterpretations of data.
  
2. Think about what quantities should be displayed.

  - Ex: Yahoo plot
  
3. Avoid pie charts in general, especially with many slices.

  - Hard to conceptualize differences in proportions.
  - Color should match the nature of the data (e.g. sequential)
  
```{r}
f_labs <- c(Borrowers = "Percent of \nall Borrowers",
            Balances = "Percent of \nall Balances")

p <- ggplot(data = studebt,
            mapping = aes(x = pct/100,
                          y = Debt,
                          fill = type))

p + geom_col() + facet_wrap(~ type, labeller = as_labeller(f_labs))
```

4. 

```{r}
library(demog)

okboomer %>% 
  filter(country == "United States") %>% 
  ggplot(aes(x = date, y = births_pct_day)) +
  geom_line(size = 0.5) +
  labs(x = "Year",
       y = "Average daily births per million")
```

```{r}
# okboomer_p <- okboomer %>% 
#   mutate(year_fct = factor(year,
#                            levels = unique(year),
#                            ordered = TRUE),
#          month_fct = factor(month,
#                             levels = rev())
```

Layer, highlight, repeat.
