---
title: "Problem Set 2"
author: "Ben Wallace"
date: "February 15, 2022"
output:
  html_document: 
    theme: cosmo
    highlight: tango
    df_print: kable
---

```{r ps03 libraries, include = FALSE}
library(tidyverse)
library(socviz)
library(slider)
library(covdata)
library(knitr)
theme_set(theme_minimal())
```

```{r ps03 knitr, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```


## Section A: Short Term Mortality Fluctuations

### Task 1

```{r belgium plot}
stmf %>% 
  filter(cname == "Belgium", sex == "b") %>% 
  group_by(year, week) %>% 
  mutate(yr_ind = year %in% 2020) %>% 
  slice(1) %>% 
  ggplot(aes(x = week, y = deaths_total, color = yr_ind, group = year)) +
  geom_line(size = 0.9) +
  scale_color_manual(values = c("gray", "red"), labels = c("2000-2019", "2020")) +
  labs(x = "Week of the Year", y = "Total Deaths", color = "Year", title = "Weekly recorded deaths in Belgium, 2000-2020") +
  theme(legend.position = "top")
  
```

### Task 2

```{r compare men women}
title <- "Death Rate by COVID-19 in Chile"
caption <- "Source: Human Mortality Database"
x_axis <- "Date"
y_axis <- "Death Rate"

df <- stmf %>% 
  filter(cname == "Chile", year == 2020, sex != "b") %>% 
  mutate(sex = ifelse(sex == "m", "Male", "Female")) %>% 
  select(rate_total, approx_date, sex)

df %>% 
  rename(Sex = sex) %>% 
  group_by(Sex) %>% 
  summarize(`Median Mortality Rate` = median(rate_total, na.rm = T)) %>% 
  knitr::kable(caption = "COVID-19 Mortality Among Chileans in 2020")

ggplot(data = df, 
       mapping = aes(x = approx_date, y = rate_total, 
                     group = sex, linetype = sex)) + 
  geom_line() +
  labs(title = title,
       caption = caption,
       linetype = "Sex",
       x = x_axis,
       y = y_axis) +
  theme(legend.position = "bottom")
```

Yes, it appears that the mortality rate in Chile was consistently higher among men then women.

### Task 3

```{r chile comparison}
df <- stmf %>% 
  filter(cname == "Chile", year == 2020, sex != "b", age_group == "75-84") %>% 
  select(death_rate, approx_date, sex) %>%
  rename(Sex = sex) %>% 
  mutate(Sex = ifelse(Sex == "m", "Male", "Female"))

df %>% 
  group_by(Sex) %>% 
  summarize(`Median Mortality Rate` = median(death_rate, na.rm = T)) %>% 
  knitr::kable(caption = "COVID-19 Mortality Among Chileans in 2020, Ages 75-84")
 
  ggplot(data = df,
         mapping = aes(x = approx_date, y = death_rate, group = Sex, linetype = Sex)) + 
    geom_line() +
    labs(title = title,
         subtitle = "Ages 75-84",
         caption = caption,
         linetype = "Sex",
         x = x_axis,
         y = y_axis) +
    theme(legend.position = "bottom")
  
```

Men in the 75-84 age group still have a higher mortality rate than women. However, the mortality rates among all Chileans in the age group higher than the national ones.

## Section B: US Deaths due to COVID-19

```{r nyt data}
nytcovstate <- nytcovstate %>% 
  group_by(state) %>% 
  mutate(daily_cases = cases - lag(cases, order_by = date), 
         daily_deaths = deaths - lag(deaths, order_by = date)) 

state_pops <- uspop %>% 
  filter(sex_id == "totsex", hisp_id == "tothisp") %>% 
  select(state_abbr, statefips, pop, state) %>% 
  rename(name = state,
         state = state_abbr, fips = statefips) %>% 
  mutate(state = replace(state, fips == "11", "DC"))
```

### Task 1

```{r left_join nytcovstate}
nytcovstate <- left_join(nytcovstate, state_pops, by = "state")
```

### Task 2

```{r death per 100,000}
nytcovstate %>%
  group_by(state) %>%
  summarize(case_rate = sum(daily_cases, na.rm = T)/100000,
            death_rate = sum(daily_deaths, na.rm = T)/100000) %>% 
  head(5) %>% 
  kable(digits = 3)
```

### Task 3

```{r 3 states}
library(lubridate)
nytcovstate %>% 
  filter(state %in% c("Wyoming", "Vermont", "Alaska")) %>% 
  ggplot(mapping = aes(x = date, y = daily_cases, group = state)) +
  geom_point(alpha = 0.8, color = "gray80") +
  labs(title = "COVID-19 in Alaska, Vermont, and Wyoming",
       x = "Date",
       y = "Daily Counts",
       caption = "Sources: New York Times, Census Bureau") +
  facet_wrap(~state)

```

### Task 4 

```{r slide_dbl states}
nytcovstate <- nytcovstate %>% 
  filter(state %in% c("Wyoming", "Vermont", "Alaska")) %>% 
  group_by(state) %>% 
  mutate(weekly_avg = slide_dbl(.x = daily_cases, .i = date, .f = ~mean(.x, na.rm = T), .before = 6, .complete = F)) 


ggplot(data = nytcovstate, mapping = aes(x = date, group = state)) +
  geom_point(mapping = aes(y = daily_cases), alpha = 0.8, color = "gray80") +
  geom_line(mapping = aes(y = weekly_avg), color = "darkred") +
  labs(title = "COVID-19 in Alaska, Vermont, and Wyoming",
       x = "Date",
       y = "Daily Cases",
       caption = "Sources: New York Times, Census Bureau") +
  facet_wrap(~ state)
```

## Section C

### Task 1 

```{r raleigh}
"Raleigh" %in% apple_mobility$region
```

### Task 2

```{r mobility}
# apple_mobility$transportation_type = factor(apple_mobility$transportation_type, levels = c("walking", "driving", "transit"))
# 
# levels(apple_mobility$transportation_type) <- c("Walking", "Driving", "Transit")

apple_mobility %>% 
  filter(region == "Raleigh") %>% 
  ggplot(mapping = aes(x = date, y = score, group = transportation_type)) +
  geom_point(alpha = 0.2, color = "gray30") +
  labs(title = "Mobility Trends in Raleigh",
       subtitle = "January 1, 2020 - October 18, 2021",
       x = NULL,
       y = "Activity Score") +
  facet_wrap(~ transportation_type)
```

### Task 3

These data roughly illustrate how "active" or "social" people in different regions are, at least for activities that require transportation. In the plot above, a reader can see both a gradual rebound in all forms of transit after the first wave of the pandemic as well as the stratification of transit forms (e.g. public transit recovers much slower than walking).

If we wanted to compare the mobility trends for two or more places, we could `facet_wrap` by region and then by mode of transportation.

```{r raleigh and nyc}

apple_mobility %>% 
  filter(region %in%  c("Raleigh", "New York City")) %>% 
  ggplot(mapping = aes(x = date, y = score, group = transportation_type)) +
  geom_point(alpha = 0.2, color = "gray30") +
  scale_y_continuous(limits  = c(0, 350)) +
  labs(title = "Mobility Trends in Raleigh and New York",
       subtitle = "January 1, 2020 - October 18, 2021",
       x = NULL,
       y = "Activity Score") + 
  facet_grid(rows = vars(transportation_type), cols = vars(region), scales = "fixed", switch = "y")
```

