---
title: "Problem Set 4"
author: "Ben Wallace"
date: "March 18, 2022"
output:
  html_document: 
    theme: cosmo
    highlight: tango
    df_print: kable
---

```{r ps4 libraries, include = F, message = F}
library(tidyverse)
library(socviz)
library(tidycensus)
library(tigris)
library(colorspace)
library(cartogram)
library(sf)
library(hrbrthemes)
```

```{r ps4 knitr, include=F}
knitr::opts_chunk$set(message = F)
```

```{r map theme, include = F}
theme_map <- function(base_size=9, base_family="") {
    require(grid)
    theme_bw(base_size=base_size, base_family=base_family) %+replace%
        theme(axis.line=element_blank(),
              axis.text=element_blank(),
              axis.ticks=element_blank(),
              axis.title=element_blank(),
              panel.background=element_blank(),
              panel.border=element_blank(),
              panel.grid=element_blank(),
              panel.spacing=unit(0, "lines"),
              plot.background=element_blank())
}
```

```{r pop names, include=F}
pop_names <- tribble(
    ~varname, ~clean_name,
    "B01003_001", "pop",
    "B01001B_001", "black",
    "B01001A_001", "white",
    "B01001H_001", "nh_white",
    "B01001I_001", "hispanic",
    "B01001D_001", "asian"
  )

inc_names <- tribble(
    ~varname, ~clean_name,
    "S1901_C01_012", "median_hh_inc")
```

```{r fips, include=F}
## Get state geography

regions <- datasets::state.name

states <- tigris::states() %>%
  filter(NAME %in% regions | NAME == "Puerto Rico" | NAME == "District of Columbia") %>% 
  shift_geometry()

## Population groups
## Code nerds note the use of `reduce` here 
fips_pop <- get_acs(geography = "county", 
                    variables = pop_names$varname, 
                    cache_table = TRUE) %>% 
  mutate(variable = reduce2(pop_names$varname, 
                            pop_names$clean_name, 
                            str_replace, 
                            .init = variable)) %>% 
  select(-moe) %>% 
  pivot_wider(names_from = variable, values_from = estimate) %>% 
  rename(fips = GEOID, name = NAME)

## Income data
fips_inc <- get_acs(geography = "county", 
                    variables = inc_names$varname,  
                    cache_table = TRUE) %>% 
  mutate(variable = str_replace(variable, 
                            inc_names$varname, 
                            inc_names$clean_name)) %>% 
  rename(fips = GEOID, name = NAME)

fips_map <- get_acs(geography = "county", 
                    variables = "B01001_001", 
                    geometry = TRUE,
                    shift_geo = FALSE,
                    cache_table = TRUE) %>% 
  shift_geometry() %>% 
  select(GEOID, NAME, geometry) %>% 
  rename(fips = GEOID, name = NAME)
```

### Task 1

```{r race ethnicity prop} 
rate_func <- function(x) {
  (x / fips_pop$pop) * 100000
}

fips_pop <- fips_pop %>% 
  mutate(across(white:hispanic, rate_func, .names = "{.col}_rate"))
```

### Task 2

```{r join pop spatial}
fips_map <- left_join(fips_map, fips_pop, by = "name")
```

I know that the merge worked because the variables in `fips_pop` have populated `fips_map`.

### Task 3

```{r black pop map}
ggplot() +
  geom_sf(data = fips_map, aes(fill = black_rate),
          color = "gray70", size = 0.1) +
  scale_fill_continuous_sequential(palette = "Blues") +
  labs(fill = "Black Americans \n per 100,000") + 
  theme_map()
```

```{r hispanic pop map}
ggplot(data = fips_map, aes(fill = hispanic_rate)) +
  geom_sf(color = "gray70", size = 0.1) +
  scale_fill_continuous_sequential(palette = "Greens") +
  labs(fill = "Hispanic Americans \n per 100,000") +
  theme_map()
```

### Task 4

```{r continuous to binned map}
fips_map <- fips_map %>% 
         mutate(across(white_rate:hispanic_rate, cut_interval, n = 5),
                across(white_rate:hispanic_rate, factor))
```

```{r black pop map binned}
values <- c("#D4E4BC", "#96ACB7", "#36558F", "#40376E", "#48233C")

levels(fips_map$black_rate) <- c("0 to 17,400", "17,400 to 34,900",
                                 "34,900 to 52,300", "52,300 to 69,800",
                                 "69,800 to 87,200")

ggplot(fips_map, aes(fill = black_rate)) +
  geom_sf(color = "gray", size = 0.1) +
  scale_fill_manual(values = values) +
  labs(fill = "Black Population \n per 100,000") + 
  theme_map()
```

```{r hispanic pop map binned}
values <- c("#8DC27C", "#E0D76E", "#DEA95F", "#CC5A4B", "#6B4B3A")

levels(fips_map$hispanic_rate) <- c("0 to 20,000", "20,000 to 40,000",
                                    "40,000 to 60,000", "60,000 to 80,000", 
                                    "80,000 to 100,000")

ggplot(fips_map, aes(fill = hispanic_rate)) +
  geom_sf(color = "gray", size = 0.1) +
  scale_fill_manual(values = values) +
  labs(fill = "Hispanic Population \n per 100,000") +
  theme_map()
```

The maps with binned racial/ethnic measures are ultimately more informative since readers have difficulty differentiating gradient values. In contrast, a reader can easily notice differences between 5 color values on the maps.

### Task 5

```{r median income}
fips_map <- left_join(fips_map, fips_inc, by = "name")

ggplot(fips_map, aes(fill = estimate)) +
  geom_sf(color = NA) +
  scale_fill_continuous_sequential(palette = "Grays", labels = scales::label_number(), breaks = seq(25000, 150000, by = 25000))+
  labs(fill = "Median Income") +
  theme_map()
  
```

### Task 6

```{r wealthiest counties}
fips_inc <- fips_inc %>% 
  arrange(desc(estimate)) %>% select(name, estimate)
  
fips_inc %>% slice_head(n = 10)
```

```{r Puerto Rico}
fips_inc %>% slice_tail(n = 5)
```


The five counties with the lowest median household income in the United States are all located in Puerto Rico. Which counties have the lowest in the mainland US?


```{r}
fips_inc %>% filter(!str_detect(name, "Puerto Rico")) %>% 
  slice_tail(n = 5)
```
